On existence of integral point sets and their diameter bounds \footnote{ This work was carried out at Voronezh State University and supported by the Russian Science Foundation grant 19-11-00197. }

### 2019-11-30 pps-linear-bound/pps-linear-lower-bound.tex
[J220] Word usage, mostly SciLexic

\begin{proof}
	Any $n^2+1$ points, including the endpoints of $\Delta$, partition the segment $\Delta$ into $n^2$
	sequential segments $\Delta_i$.
	Due to Lemma~\ref{lem:2k-1_segments}, there is at most one segment of length 1,
	at most three segments of length 2, etc.
	The following two expressions for sums conclude the proof:
	\begin{equation}
		1 + \sum_{k=1}^n (2k-1) = n^2 + 1
		,
	\end{equation}
	\begin{equation}
		\sum_{k=1}^n k(2k-1) = \frac{2}{3}n^3+\frac{1}{2}n^2-\frac{1}{6}n
		.
	\end{equation}
\end{proof}

@@

Concluding the section, we remark that
Corollary~\ref{cor:only_one_distance_1_on_straight_line}
and
Proposition~\ref{obs:estimate_points_on_straight_line}
are the only results of this section
that are used in the proofs below;
all the other are auxiliary.

@@

Basing on the exact values of $d(2,n)$ for $ 3 \leq n\leq 122$ and the estimate $d(2,123)>10000$
\cite{kurz2008bounds}, we derive the following proposition.
\begin{proposition}
	\label{obs:4_leq_n_leq_21491}
	The inequality
	\begin{equation}
		d(2,n) \geq 3^{1/4}\cdot2^{-3/2} \cdot n
	\end{equation}
	holds for $4 \leq n \leq 21491$.
\end{proposition}

@@

Performing some simple manipulations with the upper bound in Theorem~\ref{thm:varphi_k_bounds},
one can derive the following proposition.
\begin{proposition}
	\label{varphi_n_where_n_geq_21492}
	For $n \geq 21492$ we have
	\begin{equation}
		\varphi_n \leq \varphi_{n-1} \leq \frac{\beta}{\sqrt{n-2}}
		,
	\end{equation}
	where
	\begin{equation}
		\beta = \frac{1}{\sqrt{21490}} + \sqrt{ \frac{2}{\sqrt{3}} + \frac{1}{21490} } < 1.07464
		.
	\end{equation}
\end{proposition}

@@

\begin{remark}
	\label{rem:planar_set_with_minimeter_1_of_3_points}
	Applying Construction~\ref{con:planar_set_with_minimeter_1} to $k=1$ naively,
	we get $a = 3$, $I=\varnothing$, $b_\varnothing = -1$ and then
	obtain an equilateral triangle of side length 1.
	This triangle is obviously the optimal set in $\mathfrak{M}(2,3)$.
\end{remark}

@@

\begin{proof}
	Take a facher integral point set according to Construction~\ref{con:planar_set_with_minimeter_1}
	with sufficiently large height and then apply the ``blowing up'' procedure described in~\cite[Theorem 1.3]{kurz2008bounds},
	replacing the out-of-line point with an $(m-2)$-dimensional simplex of proper side length.
	That simplex will consist of $m-1$ points;
	we can throw out up to all points $M_J$, except the two points with distance 1.
	So, the set consists of $m+1$ or more points and has distance 1 occuring in it.
	As distance 1 occurs in the obtained set, it is prime.
\end{proof}

@@

For planar integral point sets containing distance 1, we can emphasise on the following problem.
\begin{problem}
	For given $n\geq 3$ and squarefree number $q \neq 1$,
	is there $M\in\mathfrak{M}(2,n)$, such that characteristic of $M$ is $q$
	and distance 1 occurs in $M$?
\end{problem}

@@

\section{Acknowledgements}
Author thanks Dr. Prof. E.M. Semenov for the fruitful discussion and ideas,
A.S. Chervinskaia for the idea of using the term ``facher'' and other linguistic advices,
and Dr. A.S. Usachev for proofreading.

>>>>

\begin{proof}
	Any $n^2+1$ points, including the endpoints of $\Delta$, split the segment $\Delta$ into $n^2$
	sequential segments $\Delta_i$.
	Due to Lemma~\ref{lem:2k-1_segments}, there is at most one segment of length 1,
	at most three segments of length 2, etc.
	The following two expressions for sums conclude the proof:
	\begin{equation}
		1 + \sum_{k=1}^n (2k-1) = n^2 + 1
		,
	\end{equation}
	\begin{equation}
		\sum_{k=1}^n k(2k-1) = \frac{2}{3}n^3+\frac{1}{2}n^2-\frac{1}{6}n
		.
	\end{equation}
\end{proof}

@@

To conclude, we remark that
Corollary~\ref{cor:only_one_distance_1_on_straight_line}
and
Proposition~\ref{obs:estimate_points_on_straight_line}
are the only results of this section
that are used in the proofs below;
all the other are auxiliary.

@@

Based on the exact values of $d(2,n)$ for $ 3 \leq n\leq 122$ and the estimate $d(2,123)>10000$
\cite{kurz2008bounds}, we derive the following proposition.
\begin{proposition}
	\label{obs:4_leq_n_leq_21491}
	The inequality
	\begin{equation}
		d(2,n) \geq 3^{1/4}\cdot2^{-3/2} \cdot n
	\end{equation}
	holds for $4 \leq n \leq 21491$.
\end{proposition}

@@

By means of some simple manipulations with the upper bound in Theorem~\ref{thm:varphi_k_bounds},
one can derive the following proposition.
\begin{proposition}
	\label{varphi_n_where_n_geq_21492}
	For $n \geq 21492$ we have
	\begin{equation}
		\varphi_n \leq \varphi_{n-1} \leq \frac{\beta}{\sqrt{n-2}}
		,
	\end{equation}
	where
	\begin{equation}
		\beta = \frac{1}{\sqrt{21490}} + \sqrt{ \frac{2}{\sqrt{3}} + \frac{1}{21490} } < 1.07464
		.
	\end{equation}
\end{proposition}

@@

\begin{remark}
	\label{rem:planar_set_with_minimeter_1_of_3_points}
	Applying Construction~\ref{con:planar_set_with_minimeter_1} to $k=1$ ad litteram,
	we get $a = 3$, $I=\varnothing$, $b_\varnothing = -1$ and then
	obtain an equilateral triangle of side length 1.
	This triangle is obviously the optimal set in $\mathfrak{M}(2,3)$.
\end{remark}

@@

\begin{proof}
	Take a facher integral point set according to Construction~\ref{con:planar_set_with_minimeter_1}
	with sufficiently large height and then apply the ``blowing up'' procedure described in~\cite[Theorem 1.3]{kurz2008bounds},
	replacing the point that is not situated on the line, with an $(m-2)$-dimensional simplex of proper side length.
	That simplex will consist of $m-1$ points;
	we can exclude practically all $M_J$ points, except the two points with distance 1.
	So, the set consists of $m+1$ or more points and has distance 1 occuring in it.
	As distance 1 occurs in the obtained set, it is prime.
\end{proof}

@@

For planar integral point sets containing distance 1, we can emphasize the following problem.
\begin{problem}
	For given $n\geq 3$ and squarefree number $q \neq 1$,
	is there $M\in\mathfrak{M}(2,n)$, such that characteristic of $M$ is $q$
	and distance 1 occurs in $M$?
\end{problem}

@@

\section{Acknowledgements}
The author thanks Dr. Prof. E.M. Semenov for the fruitful discussion and ideas,
A.S. Chervinskaia for the idea of using the term ``facher'' and proofreading,
and Dr. A.S. Usachev for proofreading.
